### 2019-05-19 tex_v5/manuscript.tex
MS ready for submission

\begin{document}
	\begin{abstract}
		The acorn weevil (\textit{Curculio} Linnaeus, 1758) rostrum (snout) exhibits remarkable flexibility and toughness derived from the microarchitecture of its exoskeleton.
		Here we characterize modifications to the composite profile of the rostral cuticle that simultaneously enhance the flexibility and toughness of the distal portion of the snout.
		Using Classical Laminate Plate Theory, we estimate the effect of these modifications on the elastic behavior of the exoskeleton.
		We show that the tensile behavior of the rostrum across six \textit{Curculio} species with high morphological variation correlates with changes in the relative layer thicknesses and orientation angles of layers in the exoskeleton.
		Accordingly, increased endocuticle thickness is strongly correlated with increased tensile strength.
		Rostrum stiffness is shown to be inversely correlated with work of fracture; thus allowing a highly curved rostrum to completely straighten without structural damage.
		Finally, we identify exocuticle rich invaginations of the occipital sutures both as a likely site of crack initiation in tensile failure, and as a source of morphological constraint on the evolution of the rostrum in \textit{Curculio} weevils.
	\end{abstract}

	%title and author block
	{\title{Exoskeletal strength and cuticle composite profile of the acorn weevil rostrum}
	\date{\today}

@@

	We additionally describe the fracture mechanics of the snout, while considering how modification of the cuticle may prevent crack formation during oviposition.
	Accordingly, the flexibility and tensile strength of the rostrum appear to be derived \emph{exclusively} from modification of the composite architecture of the exoskeleton.
	This is the first time that a modified composite profile has been reported as a means of enhancing structural elasticity in the insect exoskeleton (though see \cite{Matsumura2017}).

@@

		\subsection*{Code availability}
			R, Python, and Matlab scripts used to manipulate and analyze the raw data (as well as their outputs), produce figures, and estimate effective elastic constants are available from the corresponding author upon request.

		\subsection*{Data availability}
			Stress-strain curves for all tensile and fatigue-tested specimens are provided as PDFs (Supplement 1).
			Diagnostic plots for all PGLMMs are provided as PDFs (Supplement 2). PGLMM terms and output are provided in a PDF (Supplement 3). Raw and processed data will be provided by the corresponding author upon reasonable request.

>>>>

\begin{document}
	\begin{abstract}
		The acorn weevil (\textit{Curculio} Linnaeus, 1758) rostrum (snout) exhibits remarkable flexibility and toughness derived from the microarchitecture of its exoskeleton.
		Here we characterize modifications to the composite profile of the rostral cuticle that simultaneously enhance the flexibility and toughness of the distal portion of the snout.
		Using Classical Laminate Plate Theory, we estimate the effect of these modifications on the elastic behavior of the exoskeleton.
		We show that the tensile behavior of the rostrum across six \textit{Curculio} species with high morphological variation correlates with changes in the relative layer thicknesses and orientation angles of layers in the exoskeleton.
		Accordingly, increased endocuticle thickness is strongly correlated with increased tensile strength.
		Rostrum stiffness is shown to be inversely correlated with work of fracture; thus allowing a highly curved rostrum to completely straighten without structural damage.
		Finally, we identify exocuticle rich invaginations of the occipital sutures both as a likely site of crack initiation in tensile failure, and as a source of morphological constraint on the evolution of the rostrum in \textit{Curculio} weevils.
		We conclude that avoidance of catastrophic structural failure, as initiated in these sutures under tension, is the driving selective pressure in the evolution of the female \textit{Curculio} rostrum.
	\end{abstract}

	%title and author block
	{\title{Avoidance of catastrophic structural failure as an evolutionary constraint: Biomechanics of the acorn weevil rostrum}
	\date{\today}

@@

	Accordingly, the flexibility and tensile strength of the rostrum appear to be derived \emph{exclusively} from modification of the composite architecture of the exoskeleton.
	This is the first time that a modified composite profile has been reported as a means of enhancing structural elasticity in the insect exoskeleton (though see \cite{Matsumura2017}).
	We additionally describe the fracture mechanics of the snout, while considering how modification of the cuticle may prevent crack formation during oviposition, particularly in the fracture-prone and exocuticle-rich occipital sulci.
	Based on these findings we conclude that avoidance of catastrophic structural failure is the primary selective pressure in the evolution of the female \textit{Curculio} rostrum.

@@

		\subsection*{Code availability}
			R, Python, and Matlab scripts used to manipulate and analyze the raw data and processed data, produce figures, and estimate effective elastic constants are provided in Supplement 1.

		\subsection*{Data availability}
			Stress-strain curves for all tensile and fatigue-tested specimens are provided as PDFs (Supplement 2).
			Diagnostic plots for all PGLMMs are provided as PDFs (Supplement 3). PGLMM terms and output are provided in a PDF (Supplement 4). All measurements used for PGLMM generation are provided as a CSV file (Supplement 5). Raw and processed data will be provided by the corresponding author upon reasonable request.
